---
title: "Manuscript"
author: Ferne Kotlyar
output: pdf_document
date: "`r Sys.Date()`"
bibliography: CIEE_Course.bib
csl: Nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Introduction

According to 67 slow-motion videos (240 fps) of wild *Heliconia tortuosa* visitors, specialized hummingbirds appear to displace nectar and deposit it on the stigma as they exit the flower. This appears to occur less frequently with non-specialized hummingbirds. The goal is to compare the likelihood of nectar deposition on the stigma between specialized and non-specialized hummingbirds. To test this species-specific difference, we conducted nectar dye experiments using 3D printed bill replicas of *H. tortuosa*’s primary visitors [@betts2015; @leimberger2023], two specialized (traplining) hummingbirds: green hermits (*Phaethornis guy)*, and violet sabrewings (*Campylopterus hemileucurus)*, and two non-specialized (territorial) hummingbirds: rufous-tailed, (*Amazilia tzacatl*) and crowned woodnymphs (*Thalurania colombica*). In these dye experiments, we inject flowers with dye, use the 3D printed bills, then measure the likelihood of dye deposition on floral anthers and stigma (response = presence/absence of dye).

# Methods

**Study Location:** This study was conducted at the Las Cruces Biological Station in southern Costa Rica (Coto Brus Canton, Puntarenas Province) @stiles1975.

**Data Collection:** Open *H. tortuosa* flowers were collected from the Las Cruces Biological Station grounds. \~0.05 mL of fuchsin dye was injected into the base of the corolla tube right above the nectar chamber of *H. tortuosa* flowers @kress1983. We then inserted and extracted the bill replicas, while mimicking hummingbird behaviour, as per the slow-motion videos. Immediately after removing the bill from the flower, the anthers and stigma were placed under a field microscope (Carson MicroFlip 100x-250x LED), and the presence (or absence) of dye was recorded on the anthers and stigma separately. Each species was used 13-14 times, for a total of n = 55.

**Data Analysis:** The final results table will have a two by two factorial comparison for anthers and stigma, separately. A chi-squared and fisher exact test will be used for each table.

# Analysis

## Display Contingency Tables

```{r}
# Load packages
library(dplyr)
library(tidyr)

# Load cleaned data from the 02_outdata folder
ND_clean <- read.csv("../02_outdata/Kotlyar_CIEE_Species_Nectar_Dye_Clean.csv")

# Quick look at the data
anthers_contingency_table <- table(ND_clean$specialization, ND_clean$dye_on_anthers_binary)
stigma_contingency_table <- table(ND_clean$specialization, ND_clean$dye_on_stigma_binary)
```

**Anthers**

|                 |     |     |
|-----------------|-----|-----|
|                 | No  | Yes |
| **Territorial** | 21  | 7   |
| **Trapliner**   | 2   | 25  |

**Stigma**

|                 | No  | Yes |
|-----------------|-----|-----|
| **Territorial** | 26  | 3   |
| **Trapliner**   | 13  | 14  |

## Chi-Squared Analysis

```{r}
# Anthers
chisq.test(table(ND_clean$specialization, ND_clean$dye_on_anthers_binary))

# Stigma
chisq.test(table(ND_clean$specialization, ND_clean$dye_on_stigma_binary))
```

All results are significant, suggesting that trapliners and territorial birds deposit different amount of nectar on both the anthers and stigma of *H. tortuosa* flowers.

## Fisher Exact Test

I will also use the Fisher Exact Test due to the small count of territorials placing dye on the stigma in the contingency table.

```{r}
# Try with fishers exact test 
fisher.test(table(ND_clean$specialization, ND_clean$dye_on_anthers_binary))
fisher.test(table(ND_clean$specialization, ND_clean$dye_on_stigma_binary))
```

Once again, all results are significant, with large odd ratios. However, the confidence intervals are extremely large as well.

## Results

Although the stats were done based on hummingbird specialization, the graph will be shown based on species to include more information. The graphs below will be for the dye that lands on both the anthers and stigma based on the species. Then, we export the combined graph as a PNG into the 03_figures folder.

```{r}
# Anther Graphs

library(ggplot2)
library(ggtext)
library(stringr)

# Convert to data frame and calculate proportions
ND_clean_anthers_table <- as.data.frame.matrix(table(ND_clean$species, ND_clean$dye_on_anthers_binary)) %>%
  tibble::rownames_to_column("species") %>%
  pivot_longer(cols = c("No", "Yes"), names_to = "response", values_to = "count") %>%
  group_by(species) %>%
  mutate(prop = count / sum(count),
         ypos = rev(cumsum(rev(prop))) - prop / 2)

# Set factor
ND_clean_anthers_table$response <- factor(ND_clean_anthers_table$response, levels = c("No", "Yes"))
ND_clean_anthers_table$species <- factor(ND_clean_anthers_table$species, levels = c("CRWO", "RTAH", "GREH", "VISA"))

# Plot
ND_clean_anthers_graph <-
ggplot(ND_clean_anthers_table, aes(x = species, y = prop, fill = response, colour=species)) +
  geom_bar(stat = "identity", size = 1) +
  geom_text(aes(y = ypos, label = paste0("n=", count)), color = "white", size = 4) +
  scale_fill_manual(values = c("Yes" = "#B7B79E", "No" = "#5B4445")) +
  scale_colour_manual(values = c("CRWO" = "#1F1418", "RTAH" = "#1F1418", "GREH" = "#FF5C43", "VISA" = "#FF5C43")) +
  labs(title = "Proportion of Dye on Anthers by Species",
    x = "Species",
    y = NULL,
    fill = "Response") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 12)) + 
  guides(colour = guide_legend(override.aes = list(fill = "white")))
```

```{r}
# Stigma Graphs

# Convert to data frame and calculate proportions
ND_clean_stigma_table <- as.data.frame.matrix(table(ND_clean$species, ND_clean$dye_on_stigma_binary)) %>%
  tibble::rownames_to_column("species") %>%
  pivot_longer(cols = c("No", "Yes"), names_to = "response", values_to = "count") %>%
  group_by(species) %>%
  mutate(prop = count / sum(count),
         ypos = rev(cumsum(rev(prop))) - prop / 2)

# Set factor
ND_clean_stigma_table$response <- factor(ND_clean_stigma_table$response, levels = c("No", "Yes"))
ND_clean_stigma_table$species <- factor(ND_clean_stigma_table$species, levels = c("CRWO", "RTAH", "GREH", "VISA"))

# Pot
ND_clean_stigma_graph <-
ggplot(ND_clean_stigma_table, aes(x = species, y = prop, fill = response, colour=species)) +
  geom_bar(stat = "identity", size = 1) +
  geom_text(aes(y = ypos, label = paste0("n=", count)), color = "white", size = 4) +
  scale_fill_manual(values = c("Yes" = "#B7B79E", "No" = "#5B4445")) +
  scale_colour_manual(values = c("CRWO" = "#1F1418", "RTAH" = "#1F1418", "GREH" = "#FF5C43", "VISA" = "#FF5C43")) +
  labs(title = "Proportion of Dye on Stigma by Species",
    x = "Species",
    y = NULL,
    fill = "Response") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 12)) + 
  guides(colour = guide_legend(override.aes = list(fill = "white")))
```

```{r}
library(ggpubr)
library(gridtext)

# Figure caption
fig_caption <- str_wrap("Figure 1 | Dye deposition on anthers based on H. tortuosa's primary visitors. This includes two specialized (traplining) hummingbirds: GREH - green hermits (Phaethornis guy) and VISA - violet sabrewings, (Campylopterus hemileucurus) and two non-specialized (territorial) hummingbirds: RTAH - rufous-tailed, (Amazilia tzacatl) and CRWO - crowned woodnymphs (Thalurania colombica).")

ND_clean_combined <- ggarrange(
  ND_clean_anthers_graph + 
    labs(title = NULL) + 
    theme(plot.margin = margin(t = 20, r = 5, b = 5, l = 5)),
  ND_clean_stigma_graph + 
    labs(title = NULL) + 
    theme(plot.margin = margin(t = 20, r = 5, b = 5, l = 5)), 
  labels = c("Anthers", "Stigma"), 
  ncol = 2, 
  common.legend = TRUE, 
  legend = "bottom") 

ND_clean_combined_test <- annotate_figure(
  ND_clean_combined,
  bottom = text_grob(
    fig_caption, 
    face = "italic",
    hjust = 0.5,
    just = "left",
    size = 10))

ND_clean_combined_test

ggsave("ND_clean_combined.png", ND_clean_combined, bg="transparent", path="../03_figures", width = 8, height = 5)
```

# Discussion

In this study, we assessed how hummingbirds with different foraging strategies affect nectar deposition on *H. tortuosa*’s reproductive structures. We found that traplining (specialized) hummingbirds are significantly more likely to deposit nectar on both the anthers and stigma in comparison to territorial (non-specialized) hummingbirds.

# References
