---
title: "Manuscript"
author: Ferne Kotlyar
output: pdf_document
date: "`r Sys.Date()`"
bibliography: CIEE_Course.bib
csl: Nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Introduction

According to 67 slow-motion videos (240 fps) of wild *Heliconia tortuosa* visitors, specialized hummingbirds appear to displace nectar and deposit it on the stigma as they exit the flower. This appears to occur less frequently with non-specialized hummingbirds. The goal is to compare the likelihood of nectar deposition on the stigma between specialized and non-specialized hummingbirds. To test this species-specific difference, we conducted nectar dye experiments using 3D printed bill replicas of *H. tortuosa*’s primary visitors [@betts2015; @leimberger2023], two specialized (traplining) hummingbirds: green hermits (*Phaethornis guy)*, and violet sabrewings (*Campylopterus hemileucurus)*, and two non-specialized (territorial) hummingbirds: rufous-tailed, (*Amazilia tzacatl*) and crowned woodnymphs (*Thalurania colombica*). In these dye experiments, we inject flowers with dye, use the 3D printed bills, then measure the likelihood of dye deposition on floral anthers and stigma (response = presence/absence of dye).

# Methods

## Preliminary Power Analysis

Calculating the sample size needed for data collection.

```{r}
library(pwr)

pwr.chisq.test(
  w = 0.5,       # effect size (small=0.1, medium=0.3, large=0.5)
  df = 1,        # degrees of freedom
  sig.level = 0.05, 
  power = 0.95)

```

Assuming there will be a large effect (w=0.5), we will need a sample size of at least 52 flowers.

## Short Description

**Study Location:** This study was conducted at the Las Cruces Biological Station in southern Costa Rica (Coto Brus Canton, Puntarenas Province) @stiles1975.

**Data Collection:** Open *H. tortuosa* flowers were collected from the Las Cruces Biological Station grounds. \~0.05 mL of fuchsin dye was injected into the base of the corolla tube right above the nectar chamber of *H. tortuosa* flowers @kress1983. We then inserted and extracted the bill replicas, while mimicking hummingbird behaviour, as per the slow-motion videos. Immediately after removing the bill from the flower, the anthers and stigma were placed under a field microscope (Carson MicroFlip 100x-250x LED), and the presence (or absence) of dye was recorded on the anthers and stigma separately. Each species was used 13-14 times, for a total of n = 55.

## Data Overview

The final results table will have a two by two factorial comparison for anthers and stigma, separately, where each table will look like this:

|                 | Yes | No  |
|-----------------|-----|-----|
| **Territorial** | \-  | \-  |
| **Trapliner**   | \-  | \-  |

# Results

## Display Contingency Tables

```{r}
# Load packages
library(dplyr)
library(tidyr)

# Load cleaned data from the 02_outdata folder
ND_clean <- read.csv("../02_outdata/Kotlyar_CIEE_Species_Nectar_Dye_Clean.csv")

# Quick look at the data
print("Anthers")
table(ND_clean$specialization, ND_clean$dye_on_anthers_binary)

print("Stigma")
table(ND_clean$specialization, ND_clean$dye_on_stigma_binary)
```

## Chi-Squared Analysis

```{r}
# Anthers
chisq.test(table(ND_clean$specialization, ND_clean$dye_on_anthers_binary))

# Stigma
chisq.test(table(ND_clean$specialization, ND_clean$dye_on_stigma_binary))
```

All results are significant, suggesting that trapliners and territorial birds deposit different amount of nectar on both the anthers and stigma of *H. tortuosa* flowers.

## Fisher Exact Test

I will also use the Fisher Exact Test due to the small count of territorials placing dye on the stigma in the contingency table.

```{r}
# Try with fishers exact test 
fisher.test(table(ND_clean$specialization, ND_clean$dye_on_anthers_binary))
fisher.test(table(ND_clean$specialization, ND_clean$dye_on_stigma_binary))
```

Once again, all results are significant, with large odd ratios. However, the confidence intervals are extremely large as well.

## Plot Results

### Anthers Graph

Although the stats were done based on hummingbird specialization, the graph will be shown based on species to include more information.

```{r}
library(ggplot2)
library(ggtext)
library(stringr)

# Convert to data frame and calculate proportions
ND_clean_anthers_table <- as.data.frame.matrix(table(ND_clean$species, ND_clean$dye_on_anthers_binary)) %>%
  tibble::rownames_to_column("species") %>%
  pivot_longer(cols = c("No", "Yes"), names_to = "response", values_to = "count") %>%
  group_by(species) %>%
  mutate(prop = count / sum(count),
         ypos = rev(cumsum(rev(prop))) - prop / 2)

# Set factor
ND_clean_anthers_table$response <- factor(ND_clean_anthers_table$response, levels = c("No", "Yes"))
ND_clean_anthers_table$species <- factor(ND_clean_anthers_table$species, levels = c("CRWO", "RTAH", "GREH", "VISA"))

# Figure caption
fig_caption_anthers <- str_wrap("**Figure 1.** Dye deposition on anthers based on *H. tortuosa'*s primary visitors. This includes two specialized (traplining) hummingbirds: GREH - green hermits (*Phaethornis guy*) and VISA - violet sabrewings, (*Campylopterus hemileucurus*) and two non-specialized (territorial) hummingbirds: RTAH - rufous-tailed, (*Amazilia tzacatl*) and CRWO - crowned woodnymphs (*Thalurania colombica*).")

# Plot
ND_clean_anthers_graph <-
ggplot(ND_clean_anthers_table, aes(x = species, y = prop, fill = response, colour=species)) +
  geom_bar(stat = "identity", size = 1) +
  geom_text(aes(y = ypos, label = paste0("n=", count)), color = "white", size = 4) +
  scale_fill_manual(values = c("Yes" = "#B7B79E", "No" = "#5B4445")) +
  scale_colour_manual(values = c("CRWO" = "#1F1418", "RTAH" = "#1F1418", "GREH" = "#FF5C43", "VISA" = "#FF5C43")) +
  labs(title = "Proportion of Dye on Anthers by Species",
    x = "Species",
    y = NULL,
    fill = "Response",
    caption = str_wrap(fig_caption_anthers)) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 12),
        plot.caption = element_markdown(hjust = 0)) + 
  guides(colour = guide_legend(override.aes = list(fill = "white")))

ND_clean_anthers_graph
```

### Stigma Graph

This graph is repeated for the dye on stigma results.

```{r}
# Convert to data frame and calculate proportions
ND_clean_stigma_table <- as.data.frame.matrix(table(ND_clean$species, ND_clean$dye_on_stigma_binary)) %>%
  tibble::rownames_to_column("species") %>%
  pivot_longer(cols = c("No", "Yes"), names_to = "response", values_to = "count") %>%
  group_by(species) %>%
  mutate(prop = count / sum(count),
         ypos = rev(cumsum(rev(prop))) - prop / 2)

# Set factor
ND_clean_stigma_table$response <- factor(ND_clean_stigma_table$response, levels = c("No", "Yes"))
ND_clean_stigma_table$species <- factor(ND_clean_stigma_table$species, levels = c("CRWO", "RTAH", "GREH", "VISA"))

# Pot
ND_clean_stigma_graph <-
ggplot(ND_clean_stigma_table, aes(x = species, y = prop, fill = response, colour=species)) +
  geom_bar(stat = "identity", size = 1) +
  geom_text(aes(y = ypos, label = paste0("n=", count)), color = "white", size = 4) +
  scale_fill_manual(values = c("Yes" = "#B7B79E", "No" = "#5B4445")) +
  scale_colour_manual(values = c("CRWO" = "#1F1418", "RTAH" = "#1F1418", "GREH" = "#FF5C43", "VISA" = "#FF5C43")) +
  labs(title = "Proportion of Dye on Stigma by Species",
    x = "Species",
    y = NULL,
    fill = "Response",
    caption = expression(paste("Dye deposition on stigma based on ", italic("H. tortuosa"), "'s primary visitors. This includes two specialized (traplining) hummingbirds: GREH - green hermits (", italic("Phaethornis guy"), ") and VISA - violet sabrewings, (", italic("Campylopterus hemileucurus"), ") and two non-specialized (territorial) hummingbirds: RTAH - rufous-tailed, (", italic("Amazilia tzacatl"), ") and CRWO - crowned woodnymphs (" , italic("Thalurania colombica"), ")."))) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 12),
        plot.caption = element_text(hjust = 0)) + 
  guides(colour = guide_legend(override.aes = list(fill = "white")))

ND_clean_stigma_graph
```

### Combined Graphs

Here, we combine the graphs, then export the combined graph as a PNG into the 03_figures folder.

```{r}
library(ggpubr)

ND_clean_combined <- 
ggarrange(ND_clean_anthers_graph + labs(title = NULL) + theme(plot.margin = margin(t = 20, r = 5, b = 5, l = 5)),
          ND_clean_stigma_graph + labs(title = NULL) + theme(plot.margin = margin(t = 20, r = 5, b = 5, l = 5)), 
          labels = c("Anthers", "Stigma"), 
          ncol = 2, 
          common.legend = TRUE, 
          legend = "bottom") 

ND_clean_combined

ggsave("ND_clean_combined.png", ND_clean_combined, bg="transparent", path="../03_figures", width = 8, height = 5)
```

# References
